---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# umapr

[![Travis-CI Build Status](https://travis-ci.org/ropenscilabs/umapr.svg?branch=master)](https://travis-ci.org/ropenscilabs/umapr)

Uniform Manifold Approximation and Projection (UMAP) is a non-linear dimensionality reduction algorithm. It is similar to t-SNE but computationally more efficient. UMAP was created by Leland McInnes and John Healy ([github](https://github.com/lmcinnes/umap), [arxiv](https://arxiv.org/abs/1802.03426)). `umapr` wraps the Python implementation of UMAP to make the algorithm accessible from within R. 

## Basic use

Here is an example of running UMAP on the `iris` data set. 

```{r}
library(umapr)
# select only numeric columns
embedding <- as.data.frame(umap(as.matrix(iris[ , 1:4])))

# look at result
head(embedding)

# plot result
suppressPackageStartupMessages(library(tidyverse))
ggplot(embedding, aes(x = V1, y = V2, color = iris$Species)) + 
  geom_point()
```

`umap` returns a data frame with two columns containing the UMAP embeddings of the original data and (optionally) the original data attached.

There are a few important parameters. These are fully described in the UMAP Python [documentation](https://github.com/lmcinnes/umap/blob/bf1c3e5c89ea393c9de10bd66c5e3d9bc30588ee/notebooks/UMAP%20usage%20and%20parameters.ipynb). 

The `n_neighbor` argument can range from 2 to n-1 where n is the number of rows in the data.
 
```{r}
neighbors <- c(4, 8, 16, 32, 64, 128)

f <- lapply(neighbors, function(neighbor) {
  iris_result <- umap(as.matrix(iris[,1:4]), n_neighbors = as.integer(neighbor))
  cbind(as.data.frame(iris_result), data.frame(Species = iris$Species))
})

names(f) <- neighbors

bind_rows(f, .id = "Neighbor") %>% 
  mutate(Neighbor = as.integer(Neighbor)) %>% 
  ggplot(aes(V1, V2, color = Species)) + geom_point() + 
  facet_wrap(~ Neighbor, scales = "free")
```

The `min_dist` argument can range from 0 to 1. 

```{r}
dists <- c(0.001, 0.01, 0.05, 0.1, 0.5, 0.99)

f <- lapply(dists, function(dist) {
  iris_result <- umap(as.matrix(iris[,1:4]), min_dist = dist)
  cbind(as.data.frame(iris_result), data.frame(Species = iris$Species))
})

names(f) <- dists

bind_rows(f, .id = "Distance") %>% 
  ggplot(aes(V1, V2, color = Species)) + geom_point() + 
  facet_wrap(~ Distance, scales = "free")
```

The `distance` argument can be a bunch of stuff. 

```{r}
dists <- c("euclidean", "manhattan", "canberra", "cosine", "hamming", "dice")

f <- lapply(dists, function(dist) {
  iris_result <- umap(as.matrix(iris[,1:4]), metric = dist)
  cbind(as.data.frame(iris_result), data.frame(Species = iris$Species))
})

names(f) <- dists

bind_rows(f, .id = "Metric") %>% 
  ggplot(aes(V1, V2, color = Species)) + geom_point() + 
  facet_wrap(~ Metric, scales = "free")
```
