---
title: "UMAP in R"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Uniform Manifold Approximation and Projection (UMAP) is a non-linear dimensionality reduction algorithm. It is similar to t-SNE but computationally more efficient. UMAP was created by Leland McInnes and John Healy ([github](https://github.com/lmcinnes/umap), [arxiv](https://arxiv.org/abs/1802.03426)). `umapr` wraps the Python implementation of UMAP to make the algorithm accessible from within R. 

## Basic use

Here is an example of running UMAP on the `iris` data set. 

```{r fig.width=7}
library(umapr)
library(tidyverse)
# select only numeric columns
embedding <- umap(as.matrix(iris[ , 1:4]))

# look at result
head(embedding$returnData())

#plot the result
embedding$plot("Petal.Length")
```

`umap` returns an R6 object, which contains a data frame two columns called("UMAP1","UMAP2") containing the UMAP embeddings of the original data and the original data attached. To get the actual output as a `data.frame`, use the `$returnData()` method attached to the R6 object.

The R6 object also includes an `$explore()` method, which will bring up a Shiny app for exploring coloring the different variables on the umap plots.

```{r eval=FALSE}

embedding$explore()
```

There are a few important parameters. These are fully described in the UMAP Python [documentation](https://github.com/lmcinnes/umap/blob/bf1c3e5c89ea393c9de10bd66c5e3d9bc30588ee/notebooks/UMAP%20usage%20and%20parameters.ipynb). 

The `n_neighbor` argument can range from 2 to n-1 where n is the number of rows in the data.
 
```{r fig.width=7}
neighbors <- c(4, 8, 16, 32, 64, 128)

f <- lapply(neighbors, function(neighbor) {
  iris_result <- umap(as.matrix(iris[,1:4]), n_neighbors = as.integer(neighbor))
 
  cbind(iris_result$returnData(), Species=iris$Species)
})

names(f) <- neighbors

bind_rows(f, .id = "Neighbor") %>% 
  mutate(Neighbor = as.integer(Neighbor)) %>% 
  ggplot(aes(UMAP1, UMAP2, color = Species)) + geom_point() + 
  facet_wrap(~ Neighbor, scales = "free")
```

The `min_dist` argument can range from 0 to 1. 

```{r fig.width=7}
dists <- c(0.001, 0.01, 0.05, 0.1, 0.5, 0.99)

f <- lapply(dists, function(dist) {
  iris_result <- umap(as.matrix(iris[,1:4]), min_dist = dist)
  cbind(iris_result$returnData(), Species=iris$Species)
})

names(f) <- dists

bind_rows(f, .id = "Distance") %>% 
  ggplot(aes(UMAP1, UMAP2, color = Species)) + geom_point() + 
  facet_wrap(~ Distance, scales = "free")
```

The `distance` argument can be a bunch of stuff. 

```{r fig.width=7}
dists <- c("euclidean", "manhattan", "canberra", "cosine", "hamming", "dice")

f <- lapply(dists, function(dist) {
  iris_result <- umap(as.matrix(iris[,1:4]), metric = dist)
  cbind(iris_result$returnData(), Species=iris$Species)
})

names(f) <- dists

bind_rows(f, .id = "Metric") %>% 
  ggplot(aes(UMAP1, UMAP2, color = Species)) + geom_point() + 
  facet_wrap(~ Metric, scales = "free")
```
