---
title: "UMAP in R"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Uniform Manifold Approximation and Projection (UMAP) is a non-linear dimensionality reduction algorithm. It is similar to t-SNE but computationally more efficient. UMAP was created by Leland McInnes and John Healy ([github](https://github.com/lmcinnes/umap), [arxiv](https://arxiv.org/abs/1802.03426)). `umapr` wraps the Python implementation of UMAP to make the algorithm accessible from within R. 

## Basic use

Here is an example of running UMAP on the `iris` data set. 

```{r}
library(umapr)
# select only numeric columns
embedding <- as.data.frame(umap(as.matrix(iris[ , 1:4])))

# look at result
head(embedding)

# plot result
library(ggplot2)
ggplot(embedding, aes(x = V1, y = V2, color = iris$Species)) + 
  geom_point()
```

`umap` returns a data frame with two columns containing the UMAP embeddings of the original data and (optionally) the original data attached.

There are a few important parameters 
 
```{r}

```

